---
title: "Federated Exploration"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
# rm(list = ls())
library(MASS)
library(sn) 
library(tidyverse)
library(glmnet)

set.seed(202604)
source("CLEAR.R")
source("CLEAR-KDE.R")
source("CLEAR-KNN.R")
```


 ***CLEAR: Collaborative Local Estimation via Artificial Reference-data***


## 1. Setup and Data Generation

First, source our custom CLEAR functions and generate a federated dataset across 10 sites featuring covariate shift and varied skewness.

```{r data_generation}
# DGP (Covariate Shift with Skewness)
M <- 10                     
d <- 4               
vars <- c(paste0("X", 1:(d-1)), "Y")
n_k <- sample(200:800, M, replace = TRUE)

shuffled_sites <- sample(1:M)
size1 = size2 = round(M/3)
size3 <- M - (size1 + size2) 
profiles <- list(
  group1 = list(mu = c(50, 55, 60, 65), rho = 0.6, skew = 5, sites = shuffled_sites[1:size1]),
  group2 = list(mu = c(40, 42, 45, 48), rho = 0.4, skew = 2, sites = shuffled_sites[(size1 + 1):(size1 + size2)]),
  group3 = list(mu = c(25, 28, 30, 32), rho = 0.2, skew = 0, sites = shuffled_sites[(size1 + size2 + 1):M])
)
# 
# profiles <- list(
#   group1 = list(mu = c(0, 0, 0, 0), rho = 0.6, skew = 0, sites = shuffled_sites[1:size1]),
#   group2 = list(mu = c(0, 0, 0, 0), rho = 0.4, skew = 0, sites = shuffled_sites[(size1 + 1):(size1 + size2)]),
#   group3 = list(mu = c(0, 0, 0, 0), rho = 0.2, skew = 0, sites = shuffled_sites[(size1 + size2 + 1):M])
# )

site_data_list <- list()
variable_sd <- runif(d, min = 5, max = 20)
 
for (k in 1:M) {
  current_prof <- Filter(function(p) k %in% p$sites, profiles)[[1]]
  R <- matrix(current_prof$rho, nrow = d, ncol = d)
  diag(R) <- 1
  
  D <- diag(variable_sd[1:d])
  Sigma <- D %*% R %*% D
  
  df <- as.data.frame(rmsn( # multivariate skew-normal distribution
    n     = n_k[k], 
    # xi    = rep(current_prof$mu, length.out = d), 
    xi = current_prof$mu,
    Omega = Sigma, 
    alpha = rep(current_prof$skew, length.out = d)
  ))
  
colnames(df) <- vars
site_data_list[[k]] <- df
}

sample_size <- sapply(site_data_list, nrow); 
cat("\nSample Sizes \n"); sample_size 
total_sample_size <- sum(sapply(site_data_list, nrow))
cat("\nTotal Sample Size \n");total_sample_size
```

## 2. Execute CLEAR Pipeline

We run the CLEAR protocol to estimate local density ratios and generate the augmented artificial dataset.

```{r run_CLEAR}
# RUN CLEAR
CLEAR_results <- CLEAR(site_data_list, K_comp = 3, n_0 = 10000, inf_factor = 1.5)
# CLEAR_results <- CLEAR_KDE(site_data_list, K_comp = 3, n_0 = 10000, inf_factor = 1)
# CLEAR_results <- CLEAR_KNN(site_data_list, K_comp = 3, n_0 = 10000, inf_cfactor = 1)
cat("\n Reference Dataset Dim \n");CLEAR_results$Z_0 %>% dim()
CLEAR_results <- CLEAR_results
```

## 3. EDA Statistics Evaluation (Site 1)

Comparing the IPW estimates from the CLEAR pipeline directly against the empirical truth for Site 1.

```{r eval_stats}
target_site <- 5
truth_site1 <- calc_empirical_truth(site_data_list[[target_site]])
est_site1 <- CLEAR_results[[paste0("Site_", target_site)]]

cat(sprintf("=== EVALUATION: CLEAR vs TRUTH (Site %d) ===\n", target_site));
# Metrics (Mean, Variance, Skewness, Kurtosis)
for (stat in c("Mean", "Variance", "Skewness", "Kurtosis")) {
  cat(sprintf("\n>>> %s <<<\n", stat))
  comp_df <- data.frame(
    Variable = vars,
    Truth = round(truth_site1[[stat]], 4),
    CLEAR = round(est_site1[[stat]], 4),
    Abs_Error = round(abs(est_site1[[stat]] - truth_site1[[stat]]), 4)
  )
  print(comp_df, row.names = FALSE)
}

# Covariance Matrix
cat("\n[True Covariance]\n");print(round(truth_site1$Covariance, 2))

cat("\n[CLEAR Estimated Covariance]\n"); print(round(est_site1$Covariance, 2))
```

## 4. Visualizing Distribution Recovery

### 4.1 Local Site Recovery (Site 1)

Overlaying the CLEAR estimated distribution with the true local data.

```{r plot_local}
Z_0 <- CLEAR_results$Z_0
ipw_weights <- CLEAR_results[[paste0("Site_", target_site)]]$Weights
true_data <- site_data_list[[target_site]]

df_true <- true_data %>%
  mutate(Source = "True Local Data", 
         Weight = 1 / nrow(true_data)) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

df_CLEAR <- Z_0 %>%
  mutate(Source = "CLEAR Data", 
         Weight = ipw_weights) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

plot_df <- bind_rows(df_true, df_CLEAR)

ggplot(plot_df, aes(x = Value, weight = Weight, fill = Source, color = Source)) +
  geom_density(alpha = 0.4, adjust = 1.2, linewidth = 0.8) +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("True Local Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  scale_color_manual(values = c("True Local Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  labs(title = sprintf("Distribution Recovery: CLEAR vs True Data (Site %d)", target_site),
       x = "Variable Value", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text = element_text(face = "bold", size = 12))
```

### 4.2 Global Distribution Recovery

Evaluating the overall global population by averaging weights across all sites.

```{r plot_global}
# Calculate Global CLEAR Weights
w_global <- rep(0, nrow(Z_0))
for (h in 1:M) {
  w_global <- w_global + CLEAR_results[[paste0("Site_", h)]]$Weights
}
w_global <- w_global / M

# Prepare Global Truth Data
true_data_global <- bind_rows(site_data_list) 
df_true_global <- true_data_global %>%
  mutate(Source = "True Global Data", 
         Weight = 1 / nrow(true_data_global)) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

# Prepare Global CLEAR Data
df_CLEAR_global <- Z_0 %>%
  mutate(Source = "CLEAR Data", 
         Weight = w_global) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

plot_df_global <- bind_rows(df_true_global, df_CLEAR_global)

ggplot(plot_df_global, aes(x = Value, weight = Weight, fill = Source, color = Source)) +
  geom_density(alpha = 0.4, adjust = 1.2, linewidth = 0.8) +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  scale_color_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  labs(title = "Global Distribution Recovery: CLEAR vs Pooled True Data",
       x = "Variable Value", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text = element_text(face = "bold", size = 12))
```




## 5. Variable Selection & Downstream Modeling

Using the IPW weights to perform variable selection natively on the artificial data, comparing the discovered coefficients with those fitted on the true, private data.


```{r}
x_vars <- vars[vars != "Y"] 
X_CLEAR <- as.matrix(Z_0[, x_vars])
Y_CLEAR <- Z_0$Y
w_site1 <- CLEAR_results[[paste0("Site_", target_site)]]$Weights

# True matrices
X_true_site1 <- as.matrix(site_data_list[[target_site]][, x_vars])
Y_true_site1 <- site_data_list[[target_site]]$Y

X_true_global <- as.matrix(true_data_global[, x_vars])
Y_true_global <- true_data_global$Y

# --- Run Lasso Models ---
# Site 1 Level
cv_true_site1 <- cv.glmnet(X_true_site1, Y_true_site1, alpha = 1)
coef_true_site1 <- as.vector(coef(cv_true_site1, s = "lambda.min"))

cv_CLEAR_site1 <- cv.glmnet(X_CLEAR, Y_CLEAR, alpha = 1, weights = w_site1)
coef_CLEAR_site1 <- as.vector(coef(cv_CLEAR_site1, s = "lambda.min"))

# Global Level
cv_true_global <- cv.glmnet(X_true_global, Y_true_global, alpha = 1)
coef_true_global <- as.vector(coef(cv_true_global, s = "lambda.min"))

cv_CLEAR_global <- cv.glmnet(X_CLEAR, Y_CLEAR, alpha = 1, weights = w_global)
coef_CLEAR_global <- as.vector(coef(cv_CLEAR_global, s = "lambda.min"))

feature_labels <- c("Intercept", x_vars)

comp_site1 <- data.frame(
  Feature = feature_labels,
  Truth = round(coef_true_site1, 4),
  CLEAR = round(coef_CLEAR_site1, 4),
  Abs_Error = round(abs(coef_CLEAR_site1 - coef_true_site1), 4)
)
cat("\n>>> SITE", target_site, ": Lasso Coefficients <<<\n")
print(comp_site1, row.names = FALSE)

comp_global <- data.frame(
  Feature = feature_labels,
  Truth = round(coef_true_global, 4),
  CLEAR = round(coef_CLEAR_global, 4),
  Abs_Error = round(abs(coef_CLEAR_global - coef_true_global), 4)
)
cat("\n>>> GLOBAL POOLED: Lasso Coefficients <<<\n")
print(comp_global, row.names = FALSE)
```



# If no weight adjustment
$Z_0$ the reference data also seems well-matched with the true global data. But the value of the method is that it can recover local site distributions without sharing any IPD, and the global distribution is a byproduct of the local recoveries. The IPW weights allow us to flexibly adjust the reference data to match each site's distribution, which is crucial for accurate local recovery. 

```{r}
# Prepare Global CLEAR Data
df_CLEAR_global <- Z_0 %>%
  mutate(Source = "CLEAR Data", 
         Weight = 1) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

plot_df_global <- bind_rows(df_true_global, df_CLEAR_global)

ggplot(plot_df_global, aes(x = Value, weight = Weight, fill = Source, color = Source)) +
  geom_density(alpha = 0.4, adjust = 1.2, linewidth = 0.8) +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  scale_color_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  labs(title = "Global Distribution Recovery: CLEAR vs Pooled True Data",
       x = "Variable Value", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text = element_text(face = "bold", size = 12))

plot(density(Z_0$X3), main = "Global Density: CLEAR", ylab = "Density") # no weights
```


```{r, fig.width=10, fig.height=6}
library(ggridges)
library(patchwork)

# 1. Prepare Long-Format Weight Data
weight_data <- data.frame(
  Weight = unlist(lapply(1:M, function(h) CLEAR_results[[paste0("Site_", h)]]$Weights)),
  Site = factor(rep(1:M, each = nrow(Z_0)))
)

# 2. Site-specific vs Global Weight Distribution (Ridge Plot)
p1 <- ggplot(weight_data, aes(x = Weight, y = Site, fill = Site)) +
  geom_density_ridges(alpha = 0.6, scale = 1.5, rel_min_height = 0.01) +
  geom_vline(xintercept = 1/nrow(Z_0), linetype = "dashed", color = "red") +
  theme_minimal() +
  labs(title = "Weight Distribution by Site",
       subtitle = "Red dashed line indicates uniform weight (1/n_0)",
       x = "Normalized IPW Weight", y = "Site ID") +
  theme(legend.position = "none")

# 3. Global Weight Summary
global_weights <- data.frame(Weight = w_global)
p2 <- ggplot(global_weights, aes(x = Weight)) +
  geom_histogram(bins = 50, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(title = "Global Aggregated Weight Distribution",
       x = "Average Weight across Sites", y = "Frequency")

# 4. Weight Concentration Plot (How weights relate to a key feature, e.g., X1)
p3 <- ggplot(Z_0, aes(x = X1, y = X3, color = w_global)) +
  geom_point(alpha = 0.5, size = 0.8) +
  scale_color_viridis_c(option = "magma", name = "Global Weight") +
  theme_minimal() +
  labs(title = "Weight Concentration")

(p1 | p2) / p3
```


Remember three types of sites
```{r}
profiles$group1$sites
profiles$group2$sites
profiles$group3$sites
```


```{r, fig.width=12, fig.height=8}
M <- length(CLEAR_results) - 1 

plot_data_faceted <- lapply(1:M, function(h) {
  site_key <- paste0("Site_", h)
  Z_0 %>%
    select(X1, X3) %>%
    mutate(
      Weight = CLEAR_results[[site_key]]$Weights,
      Site = paste("Site", h)
    )
}) %>% bind_rows()

ggplot(plot_data_faceted, aes(x = X1, y = X3, color = Weight)) +
  geom_point(alpha = 0.4, size = 0.5) +
  scale_color_viridis_c(
    option = "magma", 
    name = "IPW Weight",
    trans = "sqrt" # Optional: using sqrt scale helps highlight lower weights if distribution is skewed
  ) +
  facet_wrap(~Site, ncol = 4) +
  theme_minimal(base_size = 12) +
  labs(
    title = "Weight Concentration Across Sites",
    subtitle = "Visualizing how the Reference Distribution (Z_0) is weighted to recover each local site's density") +
  theme(legend.position = "bottom",
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank())
```

If we want to say how the weights are distributed for reference data for different sites, we can also plot the weights directly:
Site 1 and 2 are similar bc they are in the same group profile, site 5 is different. 
```{r}
plot(1:50, CLEAR_results$Site_1$Weights[1:50], type = "l", main = "IPW Weights for Site 1", xlab = "Index", ylab = "Weight")
plot(1:50, CLEAR_results$Site_2$Weights[1:50], type = "l", main = "IPW Weights for Site 2", xlab = "Index", ylab = "Weight")
plot(1:50, CLEAR_results$Site_5$Weights[1:50], type = "l", main = "IPW Weights for Site 5", xlab = "Index", ylab = "Weight")
```


