---
title: "Federated Exploration"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set()
# rm(list = ls())
library(MASS)
library(sn) 
library(tidyverse)
library(glmnet)

set.seed(202604)
source("CLEAR.R")
source("CLEAR-KDE.R")
source("CLEAR-KNN.R")
```


 ***CLEAR: Collaborative Local Estimation via Artificial Reference-data***


## 1. Setup and Data Generation

First, source our custom CLEAR functions and generate a federated dataset across 10 sites featuring covariate shift and varied skewness.

```{r data_generation}
# DGP (Covariate Shift with Skewness)
M <- 10                     
d <- 4               
vars <- c(paste0("X", 1:(d-1)), "Y")
n_k <- sample(200:800, M, replace = TRUE)

shuffled_sites <- sample(1:M)
size1 = size2 = round(M/3)
size3 <- M - (size1 + size2) 
profiles <- list(
  group1 = list(mu = c(50, 55, 60, 65), rho = 0.6, skew = 30, df = 3, sites = shuffled_sites[1:size1]),
  group2 = list(mu = c(40, 42, 45, 48), rho = 0.4, skew = 20, df = 6, sites = shuffled_sites[(size1 + 1):(size1 + size2)]),
  group3 = list(mu = c(25, 28, 30, 32), rho = 0.2, skew = 10,  df = 9, sites = shuffled_sites[(size1 + size2 + 1):M])
)

site_data_list <- list()
variable_sd <- runif(d, min = 5, max = 20)
 
for (k in 1:M) {
  current_prof <- Filter(function(p) k %in% p$sites, profiles)[[1]]
  
  # Construct Sigma
  R <- matrix(current_prof$rho, nrow = d, ncol = d)
  diag(R) <- 1
  D <- diag(variable_sd[1:d])
  Sigma <- D %*% R %*% D
  
  # Use rmst (Multivariate Skew-t) instead of rmsn, heavier skewness
  df_sim <- as.data.frame(rmst(
    n     = n_k[k], 
    xi    = current_prof$mu,
    Omega = Sigma, 
    alpha = rep(current_prof$skew, length.out = d),
    nu    = current_prof$df # Degrees of freedom
  ))
  
  colnames(df_sim) <- vars
  site_data_list[[k]] <- df_sim
}

sample_size <- sapply(site_data_list, nrow); 
cat("\nSample Sizes \n"); sample_size 
total_sample_size <- sum(sapply(site_data_list, nrow))
cat("\nTotal Sample Size \n");total_sample_size
```



## 2. Execute CLEAR Pipeline

We run the CLEAR protocol to estimate local density ratios and generate the augmented artificial dataset.

```{r run_CLEAR}
# RUN CLEAR
CLEAR_results <- CLEAR(site_data_list, K_comp = 3, n_0 = 10000, inf_factor = 1)
# CLEAR_results <- CLEAR_KDE(site_data_list, K_comp = 3, n_0 = 10000, inf_factor = 1)
# CLEAR_results <- CLEAR_KNN(site_data_list, K_comp = 3, n_0 = 10000, inf_cfactor = 1)
cat("\n Reference Dataset Dim \n");CLEAR_results$Z_0 %>% dim()
CLEAR_results <- CLEAR_results
```

## 3. EDA Statistics Evaluation (Site 1)

Comparing the IPW estimates from the CLEAR pipeline directly against the empirical truth for Site 1.

```{r eval_stats}
target_site <- 1
truth_site1 <- calc_empirical_truth(site_data_list[[target_site]])
est_site1 <- CLEAR_results[[paste0("Site_", target_site)]]

cat(sprintf("=== EVALUATION: CLEAR vs TRUTH (Site %d) ===\n", target_site));
for (stat in c("Mean", "Variance")) {
  cat(sprintf("\n>>> %s <<<\n", stat))
  comp_df <- data.frame(
    Variable = vars,
    Truth = round(truth_site1[[stat]], 4),
    CLEAR = round(est_site1[[stat]], 4),
    Abs_Error = round(abs(est_site1[[stat]] - truth_site1[[stat]]), 4)
  )
  print(comp_df, row.names = FALSE)
}

cat("\n>>> Quantiles (Absolute Error: |CLEAR - Truth|) <<<\n")
quant_error <- abs(est_site1$Quantiles - truth_site1$Quantiles)
cat("\n[True Quantiles]\n");print(round(truth_site1$Quantiles, 2))
cat("\n[CLEAR Estimated Quantiles]\n"); print(round(est_site1$Quantiles, 2))


# Covariance Matrix
cov_error <- abs(est_site1$Covariance - truth_site1$Covariance)
cat("\n[True Covariance]\n");print(round(truth_site1$Covariance, 2))
cat("\n[CLEAR Estimated Covariance]\n"); print(round(est_site1$Covariance, 2))
```



## 4. Visualizing Distribution Recovery

### 4.1 Local Site Recovery (Site 1)

Overlaying the CLEAR estimated distribution with the true local data.

```{r plot_local}
Z_0 <- CLEAR_results$Z_0
ipw_weights <- CLEAR_results[[paste0("Site_", target_site)]]$Weights
true_data <- site_data_list[[target_site]]

df_true <- true_data %>%
  mutate(Source = "True Local Data", 
         Weight = 1 / nrow(true_data)) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

df_CLEAR <- Z_0 %>%
  mutate(Source = "CLEAR Data", 
         Weight = ipw_weights) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

plot_df <- bind_rows(df_true, df_CLEAR)

ggplot(plot_df, aes(x = Value, weight = Weight, fill = Source, color = Source)) +
  geom_density(alpha = 0.4, adjust = 1.2, linewidth = 0.8) +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("True Local Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  scale_color_manual(values = c("True Local Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  labs(title = sprintf("Distribution Recovery: CLEAR vs True Data (Site %d)", target_site),
       x = "Variable Value", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text = element_text(face = "bold", size = 12))
```

### 4.2 Global Distribution Recovery

Evaluating the overall global population by averaging weights across all sites.

```{r plot_global}
# Calculate Global CLEAR Weights
w_global <- rep(0, nrow(Z_0))
for (h in 1:M) {
  w_global <- w_global + CLEAR_results[[paste0("Site_", h)]]$Weights
}
w_global <- w_global / M

# Prepare Global Truth Data
true_data_global <- bind_rows(site_data_list) 
df_true_global <- true_data_global %>%
  mutate(Source = "True Global Data", 
         Weight = 1 / nrow(true_data_global)) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

# Prepare Global CLEAR Data
df_CLEAR_global <- Z_0 %>%
  mutate(Source = "CLEAR Data", 
         Weight = w_global) %>%
  pivot_longer(cols = all_of(vars), names_to = "Variable", values_to = "Value")

plot_df_global <- bind_rows(df_true_global, df_CLEAR_global)

ggplot(plot_df_global, aes(x = Value, weight = Weight, fill = Source, color = Source)) +
  geom_density(alpha = 0.4, adjust = 1.2, linewidth = 0.8) +
  facet_wrap(~Variable, scales = "free", ncol = 2) +
  theme_minimal(base_size = 14) +
  scale_fill_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  scale_color_manual(values = c("True Global Data" = "#2c7bb6", "CLEAR Data" = "#d7191c")) +
  labs(title = "Global Distribution Recovery: CLEAR vs Pooled True Data",
       x = "Variable Value", y = "Density") +
  theme(legend.position = "bottom", legend.title = element_blank(),
        strip.text = element_text(face = "bold", size = 12))
```




## 5. Variable Selection & Downstream Modeling

Using the IPW weights to perform variable selection natively on the artificial data, comparing the discovered coefficients with those fitted on the true, private data.


```{r}
x_vars <- vars[vars != "Y"] 
X_CLEAR <- as.matrix(Z_0[, x_vars])
Y_CLEAR <- Z_0$Y
w_site1 <- CLEAR_results[[paste0("Site_", target_site)]]$Weights

# True matrices
X_true_site1 <- as.matrix(site_data_list[[target_site]][, x_vars])
Y_true_site1 <- site_data_list[[target_site]]$Y

X_true_global <- as.matrix(true_data_global[, x_vars])
Y_true_global <- true_data_global$Y

# --- Run Lasso Models ---
# Site 1 Level
cv_true_site1 <- cv.glmnet(X_true_site1, Y_true_site1, alpha = 1)
coef_true_site1 <- as.vector(coef(cv_true_site1, s = "lambda.min"))

cv_CLEAR_site1 <- cv.glmnet(X_CLEAR, Y_CLEAR, alpha = 1, weights = w_site1)
coef_CLEAR_site1 <- as.vector(coef(cv_CLEAR_site1, s = "lambda.min"))

# Global Level
cv_true_global <- cv.glmnet(X_true_global, Y_true_global, alpha = 1)
coef_true_global <- as.vector(coef(cv_true_global, s = "lambda.min"))

cv_CLEAR_global <- cv.glmnet(X_CLEAR, Y_CLEAR, alpha = 1, weights = w_global)
coef_CLEAR_global <- as.vector(coef(cv_CLEAR_global, s = "lambda.min"))

feature_labels <- c("Intercept", x_vars)

comp_site1 <- data.frame(
  Feature = feature_labels,
  Truth = round(coef_true_site1, 4),
  CLEAR = round(coef_CLEAR_site1, 4),
  Abs_Error = round(abs(coef_CLEAR_site1 - coef_true_site1), 4)
)
cat("\n>>> SITE", target_site, ": Lasso Coefficients <<<\n")
print(comp_site1, row.names = FALSE)

comp_global <- data.frame(
  Feature = feature_labels,
  Truth = round(coef_true_global, 4),
  CLEAR = round(coef_CLEAR_global, 4),
  Abs_Error = round(abs(coef_CLEAR_global - coef_true_global), 4)
)
cat("\n>>> GLOBAL POOLED: Lasso Coefficients <<<\n")
print(comp_global, row.names = FALSE)
```




